Formation of $\cenn$ Molecules in Oxygen-rich Interiors of Type II Supernovae

### 2013-07-14 doc/condensation.tex
Add bins as zone properties and the effective rates docs.

In a expanding or a collapsing system the number densities change with volume. 
We'd like to study a quantity that does not depend on volume directly so that
we use the fraction of each species in number instead of number density:
\[
Y_i \equiv n_i / n_{total}
\]
where $Y_i$ is called the abundance of species i, $n_i$ is the number density
of species i and $n_{total}$ is the number density of the unit particles, here
the number density of total atoms. So $Y_i$ here indicates the number
of species i per atom. And
\[
n_{total} = n_{atom} = n_{baryon} \times Y_{atom} 
\]
where $n_{baryon} = \rho N_A$ is the baryon number density and $Y_{atom}$ is
the atom fraction, i.e. number of atoms per baryon. Take pure $^{12}$C for 
example, $Y_{atom} = 1 / 12$. So Eq.\ref{eq:number_density} can be rewrite as
\begin{equation}
\frac{{\rm d}Y_1}{{\rm d}t} = 
  -k n_{atom} =
  -k \rho N_A Y_{atom} Y_1 Y_2 + \lambda_\gamma Y_3
\label{eq:abundance}
\end{equation}
where $\rho$ is the mass density and $N_A$ is the Avogadro number. Then what
we need to study is how the abundances of each species change with time.

@@

The abundance change rates of $P_{begin}$, $P_{end}$ and $P_1$ due to 
reaction $P_{begin} + X P_1 \longleftrightarrow P_{end} + \gamma$ are
\[
\frac{{\rm d}Y_{begin}}{{\rm d}t} =
  -k^{eff} Y_{atom} Y_{begin} Y_1 + \lambda_\gamma^{eff} Y_{end}  
\]
\[
\frac{{\rm d}Y_{end}}{{\rm d}t} =
  k^{eff} Y_{atom} Y_{begin} Y_1 - \lambda_\gamma^{eff} Y_{end}  
\]
\[
\frac{{\rm d}Y_1}{{\rm d}t} =
  X( -k^{eff} Y_{atom} Y_{begin} Y_1 + \lambda_\gamma^{eff} Y_{end} )
\]
where $\lambda_\gamma^{eff}$ is the effective photodissociation rate which
will be discussed later and $X$ is the factor to be determined that ensure
the total atom number conserved. Atom conservation requires
\[
N_{begin} \times \frac{{\rm d}Y_{begin}}{{\rm d}t} +
N_{end} \times \frac{{\rm d}Y_{end}}{{\rm d}t} +
\frac{{\rm d}Y_1}{{\rm d}t} = 0.
\] 
So we have
\[
X = N_{end} - N_{begin}
\]

>>>>

In a expanding or a collapsing system the number densities change with volume. 
We'd like to study a quantity that does not depend on volume directly so that
we use the fraction of each species in number instead of number density:
\[
Y_i \equiv n_i / n_{total}
\]
where $Y_i$ is called the abundance of species i, $n_i$ is the number density
of species i and $n_{total}$ is the number density of the unit particles, here
the number density of total atoms. So $Y_i$ here indicates the number
of species i per atom. And
\[
n_{total} = n_{atom} = n_{baryon} \times Y_{atom} 
\]
where $n_{baryon} = \rho N_A$ is the baryon number density and $Y_{atom}$ is
the atom fraction, i.e. number of atoms per baryon. Take pure $^{12}$C for 
example, $Y_{atom} = 1 / 12$. So Eq.\ref{eq:number_density} can be rewrite as
\begin{equation}
\frac{{\rm d}Y_1}{{\rm d}t} = 
  -k n_{atom} Y_1 Y_2 + \lambda_\gamma Y_3 =
  -k \rho N_A Y_{atom} Y_1 Y_2 + \lambda_\gamma Y_3
\label{eq:abundance}
\end{equation}
where $\rho$ is the mass density and $N_A$ is the Avogadro number. Then what
we need to study is how the abundances of each species change with time.

@@

The abundance change rates of $P_{begin}$, $P_{end}$ and $P_1$ due to 
reaction $P_{begin} + X P_1 \longleftrightarrow P_{end} + \gamma$ are
\[
\frac{{\rm d}Y_{begin}}{{\rm d}t} =
  -k^{eff} n_{atom} Y_{begin} Y_1 + \lambda_\gamma^{eff} Y_{end}  
\]
\[
\frac{{\rm d}Y_{end}}{{\rm d}t} =
  k^{eff} n_{atom} Y_{begin} Y_1 - \lambda_\gamma^{eff} Y_{end}  
\]
\[
\frac{{\rm d}Y_1}{{\rm d}t} =
  X( -k^{eff} n_{atom} Y_{begin} Y_1 + \lambda_\gamma^{eff} Y_{end} )
\]
where $\lambda_\gamma^{eff}$ is the effective photodissociation rate which
will be discussed later and $X$ is the factor to be determined that ensure
the total atom number conserved. Atom conservation requires
\[
N_{begin} \times \frac{{\rm d}Y_{begin}}{{\rm d}t} +
N_{end} \times \frac{{\rm d}Y_{end}}{{\rm d}t} +
\frac{{\rm d}Y_1}{{\rm d}t} = 0.
\] 
So we have
\[
X = N_{end} - N_{begin}
\]

\section{Network Matrix}
Let's look at the reaction between the last species in the continuous network
and the first bin, and neglect the photodissociation here:
\[
P_{last} + XP_1 \to P_{b1}
\]
The abundance change rates are then (using $n_a$ to denote $n_{atom}$)
\[
\frac{{\rm d}Y_{last}}{{\rm d}t} =
  -k^{eff}_{last} \, n_a Y_{last} Y_1
\]
\[
\frac{{\rm d}Y_{b1}}{{\rm d}t} =
  k^{eff}_{last} \, n_a Y_{last} Y_1
\]
\[
\frac{{\rm d}Y_1}{{\rm d}t} =
  ( -k^{eff}_{last} \, n_a Y_{last} Y_1 )(N_{b1} - N_{last})
\]
Then the changes of the root-finding functions are (how to call them?)
\[
f_{last} \: +\!\!= \: k^{eff}_{last} \, n_a Y_{last} Y_1
\]
\[
f_{1} \: +\!\!= \: (k^{eff}_{last} \, n_a Y_{last} Y_1) (N_{b1} - N_{last})
\]
To find the abundance changes in the continuous network, we need to solve the
matrix
\[
AX=B
\]
The matrix element changes due to this reaction are
\[
A_{last,last} \: +\!\!= \: k^{eff}_{last} \, n_a Y_1
\]
\[
A_{last,1} \: +\!\!= \: k^{eff}_{last} \, n_a Y_{last}
\]
\[
A_{1,last} \: +\!\!= \: (k^{eff}_{last} \, n_a Y_1)(N_{b1} - N_{last})
\]
\[
A_{1,1} \: +\!\!= \: (k^{eff}_{last} \, n_a Y_{last})(N_{b1} - N_{last})
\]
The right-hand-side vector changes are
\[
B_{last} \: +\!\!= \: -f_{last} = -k^{eff}_{last} \, n_a Y_{last} Y_1 
\]
\[
B_1 \:+\!\!=\: -f_1 = -(k^{eff}_{last}\,n_a Y_{last} Y_1)(N_{b1}-N_{last})
\]

Now let's look at the reactions between bins. $P_{bi} + XP_1 \to P_{b(i+1)}$,
where $i$ goes from 1 to $N-1$ with $N$ as the number of bins.
\[
\left( \frac{{\rm d}Y_1}{{\rm d}t}\right) _i =
  ( -k^{eff}_{bi} \, n_a Y_{bi} Y_1 )(N_{b(i+1)} - N_{bi})
\]
\[
(f_{1})_i \: +\!\!= \: 
  (k^{eff}_{bi} \, n_a Y_{bi} Y_1) (N_{b(i+1)} - N_{bi})
\]
\[
(A_{1,1})_i \: +\!\!= \: 
  (k^{eff}_{bi} \, n_a Y_{bi}) (N_{b(i+1)} - N_{bi})
\]
\[
(B_1)_i \:+\!\!=\: -(f_1)_i = 
  -(k^{eff}_{bi} \, n_a Y_{bi} Y_1) (N_{b(i+1)} - N_{bi})
\]
We first solve the matrix equation for current $Y_{bi}$, and then compute
the abundance changes in bins using the new abundances in the network.
Then feed the new bin abundances back to the matrix. Iterate until converge.

\section{Bin Abundances Change}

Now let's look at the abundance changes in bins for each reaction discussed
above. We are going to solve these changes implicitly. For $Y_{b1}$,
\[
\frac {Y_{b1}(t+\Delta t) - Y_{b1}(t)} {\Delta t} =
  k^{eff}_{last} \, n_a Y_{last}(t+\Delta t) Y_1(t+\Delta t) -
  k^{eff}_{b1} \, n_a Y_{b1}(t+\Delta t) Y_1(t+\Delta t)
\]
\[
Y_{b1}(t+\Delta t) = \frac
  {k^{eff}_{last} \, n_a Y_{last}(t+\Delta t) Y_1(t+\Delta t) \Delta t +
    Y_{b1}(t)}
  {1 + k^{eff}_{b1} \, n_a Y_1(t+\Delta t) \Delta t}
\]
Similarly for $Y_{bi}$ as $i$ goes from 2 to $N-1$:
\[
Y_{bi}(t+\Delta t) = \frac
  {k^{eff}_{b(i-1)} \, n_a Y_{b(i-1)}(t+\Delta t) Y_1(t+\Delta t) \Delta t +
    Y_{bi}(t)}
  {1 + k^{eff}_{bi} \, n_a Y_1(t+\Delta t) \Delta t}
\]
For $Y_{bN}$ we don't have any destructive terms for now,
\[
Y_{bN}(t+\Delta t) = 
  k^{eff}_{b(N-1)} \, n_a Y_{b(N-1)}(t+\Delta t) Y_1(t+\Delta t) \Delta t +
    Y_{bi}(t)
\]
